---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
## Global Plastic Production Forecasting and Time Series Analysis

**Overview**

In this project, we'll analyze the time series and build a prediction model to predict the Global Plastic Production for the next 10 years.

we'll be doing the following steps to implement our project:

* We'll collect the data
* we'll explore the data by checking for missing values, outliers, irrelevant data points and irrelevant columns
* We'll scale the data so that all the values are in the range of 1-100
* Then we'll convert the data into a time series and split the data into train and test set
* We'll build our prediction model using the Holt's Exponential Smoothing method

**Data**

* The dataset Global Plastic Production has been taken from Our world in data
* The data available is from the year 1950 to 2015

#### Libraries required

```{r}
library(dplyr)
library(tseries)
library(forecast)
library(scales)
library(ggplot2)
library(expsmooth)
```


##### Loading the data
```{r}
data <- read.csv("C:/Users/user/Desktop/global-plastics-production.csv")
head(data)
```
##### about the data
```{r}
str(data)
summary(data)
```
##### Checking for missing values
```{r}
any(is.na(data))
```
#### removing the non-relevant columns
```{r}
df <- select(data, -c(Entity, Code))
str(df)
```
#### renaming the columns
```{r}
colnames(df) = c("Year", "Gpp")
```

##### Scaling the data so that it's in the range of 1-100, this will simplify to analyse and understand the computation
```{r}
df$Gpp <- rescale(df$Gpp, to = c(1,100))
head(df)
df$Gpp
```

#### Converting the data into a time series
```{r}
tsdata <- ts(df$Gpp, frequency = 1, start = 1950, end = 2015)
plot(aggregate(tsdata, FUN = mean))
```

#### Splitting the data into training set and test set
```{r}
trainingdata <- ts(df$Gpp, frequency = 1, start = 1950, end = 2005)
testdata <- tail(df, 10)
```

#### Plotting the data, from the plot we can observe that the data follows an exponential trend
```{r}
plot(tsdata)
```

#### Building the prediction model using the Holt's Exponential Smoothing method and we are training our model using the train dataset
```{r}
model <- holt(trainingdata, h = 10)
summary(model)
```
#### plotting the data with predictions on the training data
```{r}
autoplot(model)
```

#### Predicted values
```{r}
pred = forecast(model)
pred$mean
```
#### Accuracy of the model predictions
```{r}
accuracy(pred$mean, testdata$Gpp)
```

#### residuals
```{r}
checkresiduals(model)
```


#### Prediction from 2015 for the next 10 years, till 2025
```{r}
fdata <- ts(df$Gpp, frequency = 1, start = 1950, end = 2015)
ft <- holt(fdata, h = 15)
summary(ft)
autoplot(ft)
```

#### An RMSE value of 2.4 indicates an accuracy of 97% in our predictions.

#### According to our predictions the global plastic production is expected to double in the next 10 years.
